---
title: "01_deduplicate"
author: "Pipeline Agent"
date: "`r Sys.Date()`"
output: html_document
---

### Objective
This notebook serves as the gatekeeper for the analysis pipeline. It systematically identifies and resolves duplicate publications from the raw data, ensuring that each unique study is represented only once in all downstream analyses. The entire process is designed for full auditability, generating a series of log files that provide a transparent record of every decision made.

**Robustness Note**: This script is designed to run successfully even if there are no input files. In that scenario, it will produce empty log files and empty deduplicated datasets.

```{r dedup_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# This script assumes 00_setup.Rmd has been run and its objects are available.
```

### 1. Prepare Master Publication Registry
First, we create a comprehensive registry of all unique publications found across the datasets. This registry forms the basis for the deduplication process. We assign a unique internal ID (`registry_id`) to each row for stable referencing.

```{r create_registry}
# Combine metadata from all relevant datasets to build a master list
# We prioritize the dedicated study_metadata file
study_ids_from_all_data <- unlist(lapply(raw_data, function(df) if("study_id" %in% names(df)) df$study_id else NULL)) %>%
  unique() %>%
  na.omit()

# Create a base tibble of all unique study_ids
master_registry <- tibble(study_id = study_ids_from_all_data)

# Join with the main metadata file
if (nrow(raw_data$study_metadata) > 0) {
  master_registry <- master_registry %>%
    left_join(raw_data$study_metadata, by = "study_id")
} else {
  # If metadata file is empty, create placeholder columns
  master_registry <- master_registry %>%
    mutate(doi = NA_character_, pmid = NA_character_, title = NA_character_, 
           authors = NA_character_, publication_year = NA_character_, journal = NA_character_)
}

# Add a stable internal ID and normalize key fields for matching
# Ensure placeholder columns exist even if the join fails or metadata is incomplete
expected_cols <- c("doi" = NA_character_, "pmid" = NA_character_, "title" = NA_character_, 
                   "authors" = NA_character_, "publication_year" = NA_character_, "journal" = NA_character_)
master_registry <- master_registry %>%
  add_column(!!!expected_cols[setdiff(names(expected_cols), names(.))])

master_registry <- master_registry %>%
  mutate(
    registry_id = row_number(),
    norm_title = normalize_text(title),
    norm_authors = normalize_authors(authors),
    pub_year = parse_year(publication_year)
  ) %>%
  select(registry_id, study_id, doi, pmid, title, authors, pub_year, norm_title, norm_authors, everything())

# Handle case where master_registry is empty
if (nrow(master_registry) == 0) {
  message("Master registry is empty. No data to deduplicate.")
  # Create empty tibbles for all outputs to ensure downstream files can be created
  deduplicated_data <- setNames(lapply(raw_data, function(df) df[0,]), names(raw_data))
  deduplication_map <- tibble(registry_id = integer(0), study_id = character(0), cluster_id = integer(0), status = character(0), primary_study_id = character(0))
}

# Save the initial registry for audit purposes
if(nrow(master_registry) > 0) {
    write_csv(master_registry, file.path(config$paths$dedup_dir, "01_master_registry.csv"))
    message("Saved master publication registry.")
} else {
    write_csv(tibble(), file.path(config$paths$dedup_dir, "01_master_registry.csv"))
    message("Saved empty master publication registry.")
}

```

### 2. Identify Candidate Duplicate Pairs
We use a two-pronged approach to find potential duplicates.

#### 2.1. Exact Matching
We identify pairs that share the same DOI or PMID. These are considered definite duplicates.

```{r exact_matching}
if (nrow(master_registry) > 0) {
  # Find pairs with the same DOI (and the DOI is not NA)
  doi_pairs <- master_registry %>%
    filter(!is.na(doi)) %>%
    group_by(doi) %>%
    filter(n() > 1) %>%
    summarise(pairs = list(t(combn(registry_id, 2))), .groups = 'drop') %>%
    unnest_wider(pairs, names_sep = "_") %>%
    select(id1 = pairs_1, id2 = pairs_2) %>%
    mutate(match_type = "exact_doi")

  # Find pairs with the same PMID (and the PMID is not NA)
  pmid_pairs <- master_registry %>%
    filter(!is.na(pmid)) %>%
    group_by(pmid) %>%
    filter(n() > 1) %>%
    summarise(pairs = list(t(combn(registry_id, 2))), .groups = 'drop') %>%
    unnest_wider(pairs, names_sep = "_") %>%
    select(id1 = pairs_1, id2 = pairs_2) %>%
    mutate(match_type = "exact_pmid")

  exact_pairs <- bind_rows(doi_pairs, pmid_pairs) %>%
    distinct()
} else {
  exact_pairs <- tibble(id1 = integer(0), id2 = integer(0), match_type = character(0))
}
```

#### 2.2. Fuzzy Matching
We create a grid of all possible pairs and score them based on title, author, and year similarity.

```{r fuzzy_matching}
if (nrow(master_registry) > 1) {
  # Create all possible pairs of registry IDs
  all_possible_pairs <- t(combn(master_registry$registry_id, 2)) %>%
    as_tibble(.name_repair = ~ c("id1", "id2"))

  # Join with the registry to get the data for each pair
  pairs_to_score <- all_possible_pairs %>%
    left_join(master_registry, by = c("id1" = "registry_id")) %>%
    left_join(master_registry, by = c("id2" = "registry_id"), suffix = c("_1", "_2"))

  # Calculate similarity scores
  scored_pairs <- pairs_to_score %>%
    mutate(
      title_dist = mapply(calculate_levenshtein, norm_title_1, norm_title_2),
      author_jaccard = mapply(calculate_jaccard, norm_authors_1, norm_authors_2),
      year_diff = abs(pub_year_1 - pub_year_2)
    )

  # Filter for pairs that meet the fuzzy matching criteria from the config
  fuzzy_pairs <- scored_pairs %>%
    filter(
      title_dist <= config$deduplication$title_levenshtein_dist,
      author_jaccard >= config$deduplication$author_jaccard_min,
      year_diff <= config$deduplication$year_diff_max
    ) %>%
    select(id1, id2) %>%
    mutate(match_type = "fuzzy")
} else {
  scored_pairs <- tibble() # For audit log
  fuzzy_pairs <- tibble(id1 = integer(0), id2 = integer(0), match_type = character(0))
}

# Combine all candidate pairs
candidate_pairs <- bind_rows(exact_pairs, fuzzy_pairs) %>%
  distinct(id1, id2, .keep_all = TRUE)

# Save the scored pairs and candidate pairs for audit
if(exists("scored_pairs") && nrow(scored_pairs) > 0) {
    write_csv(scored_pairs %>% select(id1, id2, title_dist, author_jaccard, year_diff), file.path(config$paths$dedup_dir, "02a_all_scored_pairs.csv"))
} else {
    write_csv(tibble(), file.path(config$paths$dedup_dir, "02a_all_scored_pairs.csv"))
}
write_csv(candidate_pairs, file.path(config$paths$dedup_dir, "02b_candidate_pairs.csv"))
message("Identified and saved candidate duplicate pairs.")
```

### 3. Resolve Duplicate Clusters
Using the `igraph` package, we treat the pairs as edges in a graph to find connected components, which represent clusters of duplicates.

```{r resolve_clusters}
if (nrow(candidate_pairs) > 0) {
  # Create a graph from the pairs
  g <- graph_from_data_frame(candidate_pairs[, c("id1", "id2")], directed = FALSE)

  # Decompose the graph into clusters
  clusters <- decompose(g)

  # Assign a cluster_id to each record
  cluster_map <- tibble(
    registry_id = as.integer(names(V(g))),
    cluster_id = components(g)$membership
  )

  # Create a summary of the clusters
  duplicate_clusters <- master_registry %>%
    left_join(cluster_map, by = "registry_id") %>%
    # Records not in any pair get their own cluster_id
    mutate(cluster_id = ifelse(is.na(cluster_id), registry_id + max(cluster_map$cluster_id, na.rm = TRUE), cluster_id)) %>%
    arrange(cluster_id, registry_id)
} else {
  # If no pairs, every record is its own cluster
  duplicate_clusters <- master_registry %>%
    mutate(cluster_id = registry_id)
}

write_csv(duplicate_clusters, file.path(config$paths$dedup_dir, "03_duplicate_clusters.csv"))
message("Resolved and saved duplicate clusters.")
```

### 4. Select Primary Record per Cluster
For each cluster, we select one "primary" record based on the strict hierarchy defined in `config.yml`. All other records in the cluster will be marked for removal.

```{r select_primary}
if (nrow(duplicate_clusters) > 0) {
  # Add fields needed for the selection hierarchy
  selection_data <- duplicate_clusters %>%
    mutate(
      has_doi = !is.na(doi),
      has_pmid = !is.na(pmid),
      # Count non-NA values in key fields
      completeness_score = rowSums(!is.na(select(., study_id, title, authors, pub_year, journal)))
    )

  # Apply the selection hierarchy
  primary_records <- selection_data %>%
    group_by(cluster_id) %>%
    arrange(
      # The hierarchy is applied in reverse order of preference due to how arrange works
      desc(has_doi),
      desc(has_pmid),
      desc(completeness_score),
      desc(pub_year),
      registry_id # Tie-breaker: take the first one encountered
    ) %>%
    slice(1) %>%
    ungroup()

  # Create the final deduplication map
  deduplication_map <- duplicate_clusters %>%
    mutate(
      status = ifelse(registry_id %in% primary_records$registry_id, "primary", "duplicate"),
      primary_study_id = primary_records$study_id[match(cluster_id, primary_records$cluster_id)]
    )
} else {
  deduplication_map <- tibble(registry_id = integer(0), study_id = character(0), cluster_id = integer(0), status = character(0), primary_study_id = character(0))
}

write_csv(deduplication_map, file.path(config$paths$dedup_dir, "04_deduplication_map.csv"))
message("Selected primary records and saved the final deduplication map.")
```

### 5. Apply Deduplication to All Datasets
Finally, we use the `deduplication_map` to filter all input datasets, keeping only the rows that correspond to a primary record.

```{r apply_deduplication}
if (nrow(deduplication_map) > 0) {
  # Get the list of study_ids to keep
  ids_to_keep <- deduplication_map %>%
    filter(status == "primary") %>%
    pull(study_id)

  # Filter each dataset in the raw_data list
  deduplicated_data <- lapply(raw_data, function(df) {
    if ("study_id" %in% names(df)) {
      df %>% filter(study_id %in% ids_to_keep)
    } else {
      df # Return dataframes without study_id as is
    }
  })
} else {
  # If no data, the deduplicated data is the same as the raw (empty) data
  deduplicated_data <- raw_data
}

message("Deduplication applied to all datasets. The 'deduplicated_data' list is now available.")
```

### Deduplication Complete
The process is finished. The `deduplicated_data` list contains clean datasets ready for analysis. The `analysis/results_r/dedup/` directory contains a full audit trail.
